Automatic Amortized Analysis for Imperative Code

### 2014-03-10 main.tex
start soundness proof

\usepackage[utf8]{inputenc} %for utf8 input
\usepackage{amssymb} %for shift symbol
\usepackage{amsmath}
\usepackage{listings} %for code
\usepackage{mathpartir} %for typing rules
\usepackage{microtype} %better micro typing
\usepackage{stmaryrd} %for llbracket
\usepackage{mathabx} % for boxes
\usepackage{graphicx} %to include png images
\usepackage{xcolor} %for colors
\usepackage{url}

@@

E := x + y \mid x - y % \mid y
\end{mathpar}
%
We define a deductive system to derive judgements of the
form $\Gamma; Q \vdash S \dashv Q'; \Gamma'$ with the following
intended semantics
$$
\forall H\, H'\, S\, n,
    ( H \models \Gamma \land
      (H, S) \Downarrow_n H'
    ) \Rightarrow
    ( H' \models \Gamma' \land
      \Phi_{H}(Q) \ge n + \Phi_{H'}(Q')
    ).
$$

@@

\begin{theorem}[Terminating case]
For any $S$, $H$, $H'$, $q$, $q'$, $\Gamma$, $\Gamma'$, $Q$ and $Q'$ such
that $\Gamma; Q \vdash S \dashv Q'; \Gamma'$ and $H \models \Gamma$ and
$(H,s) \Downarrow_{(q,q')} H'$, we have
$H' \models \Gamma'$ and
$q \le \Phi_H(Q)$ and
$\Phi_H(Q) \ge q' + \Phi_{H'}(Q')$.
\end{theorem}

>>>>

\usepackage[utf8]{inputenc} %for utf8 input
\usepackage{amssymb} %for shift symbol
\usepackage{amsmath}
\usepackage{listings} %for code
\usepackage{mathpartir} %for typing rules
\usepackage{microtype} %better micro typing
\usepackage{stmaryrd} %for llbracket
\usepackage{mathabx} % for boxes
\usepackage{graphicx} %to include png images
\usepackage{xcolor} %for colors
\usepackage{url}
\usepackage{enumitem}

@@

E := x + y \mid x - y % \mid y
\end{mathpar}
%
We define a deductive system to derive judgements of the
form $\Gamma; Q \vdash S \dashv Q'; \Gamma'$.  The semantics
of this judgement are defined below in the soundness section.

@@

\begin{theorem}[Terminating case]
For any $S$, $H$, $H'$, $q$, $q'$, $\Gamma$, $\Gamma'$, $Q$ and $Q'$ such
that $\Gamma; Q \vdash S \dashv Q'; \Gamma'$ and $H \models \Gamma$ and
$(H,s) \Downarrow_{(q,q')} H'$, we have
$H' \models \Gamma'$ and
$q \le \Phi_H(Q)$ and
$\Phi_H(Q) - \Phi_{H'}(Q') \ge q - q'$.
\end{theorem}

\begin{proof} By induction on the big step semantics.

\begin{itemize}

\item Case {\sc S:Skip}.
  We have $(H, \code{skip}) \Downarrow_{M_s} H$.
  By assumption, $H \models \Gamma$. \\
  Since $\Phi_H(Q + M_s) = \Phi_H(Q) + M_s$,
  \begin{itemize}[topsep=0pt]
  \item
    if $M_s \ge 0$ we have,
      $M_s \le \Phi_H(Q + M_s)$ and
      $\Phi_H(Q + M_s) - \Phi_H(Q) \ge M_s - 0$;
  \item
    if $M_s < 0$ we have,
      $0 \le \Phi_H(Q + M_s)$ and
      $\Phi_H(Q + M_s) - \Phi_H(Q) \ge 0 - (-M_s)$.
  \end{itemize}

\item Case {\sc S:Seq}.
  By transitivity and induction hypothesis, it is
  easy to get $H'' \models \Gamma''$.  Now we prove
  the quantitative part of the theorem.  We have
  $(q, q') = M_s^1 \cdot (q_1, q'_1) \cdot M_s^2 \cdot (q_2, q'_2)$.
  We assume $\forall i \in \{1, 2\},~M_s^i \ge 0$.
  \begin{itemize}[topsep=0pt]
  \item
    The first case is when $q'_1 > M_s^2 + q_2$ then $q = M_s^1 + q_1$
    and since $q_1 \le \Phi_H(Q - M_s^1)$ (by induction) we have
    $q \le \Phi_H(Q)$.
  \item
    The second case is when $q'_1 \le M_s^2 + q_2$ then $q = M_s^1 + q_1 + M_s^2 + q_2 - q'_1$. \\
    Since $\Phi_H(Q - M_s^1) - \Phi_{H'}(Q') \ge q_1 - q'_1$
    and $q_2 \le \Phi_{H'}(Q' - M_s^2)$ we have:
    \begin{align*}
    q &\le M_s^1 + \Phi_H(Q - M_s^1) - \Phi_{H'}(Q') + M_s^2 + q_2 \\
      &= \Phi_H(Q) + (q_2 - \Phi_{H'}(Q' - M_s^2)) \le \Phi_H(Q).
   \end{align*}
  \end{itemize}
  We obtain $\Phi_H(Q) - \Phi_{H''}(Q'') \ge q - q'$ by simply chaining
  the two induction hypothesis and using the fact that $q - q' =
  (q_1 + M_s^1 - q'_1) + (q_2 + M_s^2 - q'_2)$.

\item Case {\sc S:WhileStop}.
\item Case {\sc S:WhileStep}.

\end{itemize}
\end{proof}
